---
title: "Practical Machine Learning—Course Project"
output:
  md_document:
    toc: yes
    variant: markdown_github
  html_document:
    toc: yes
  pdf_document:
    highlight: tango
    number_sections: yes
    toc: no
---

#Summary
The purpose of this study is to identify the parameters of a set of motions that can be used to identity a particular activity.

The data was provided by [http://groupware.les.inf.puc-rio.br/har](http://groupware.les.inf.puc-rio.br/har)

#Initiate Libraries

```{r , echo=TRUE , warning=FALSE,message=FALSE}
#initialize library

library(caret,quietly=TRUE,warn.conflicts=FALSE)
require(caret,quietly=TRUE,warn.conflicts=FALSE)

library(AppliedPredictiveModeling,quietly=TRUE,warn.conflicts=FALSE)
require(AppliedPredictiveModeling,quietly=TRUE,warn.conflicts=FALSE)

library(ElemStatLearn,quietly=TRUE,warn.conflicts=FALSE)
require(ElemStatLearn,quietly=TRUE,warn.conflicts=FALSE)

library(pgmm,quietly=TRUE,warn.conflicts=FALSE)
require(pgmm,quietly=TRUE,warn.conflicts=FALSE)

library(rpart,quietly=TRUE,warn.conflicts=FALSE)
require(rpart,quietly=TRUE,warn.conflicts=FALSE)

library(gbm,quietly=TRUE,warn.conflicts=FALSE)
require(gbm,quietly=TRUE,warn.conflicts=FALSE)

library(lubridate,quietly=TRUE,warn.conflicts=FALSE)
require(lubridate,quietly=TRUE,warn.conflicts=FALSE)

library(forecast,quietly=TRUE,warn.conflicts=FALSE)
require(forecast,quietly=TRUE,warn.conflicts=FALSE)

library(e1071,quietly=TRUE,warn.conflicts=FALSE)
require(e1071,quietly=TRUE,warn.conflicts=FALSE)

library(doBy,quietly=TRUE,warn.conflicts=FALSE)
require(doBy,quietly=TRUE,warn.conflicts=FALSE)
```

# Data

The data was sourced from the following locations

* Training Data

[https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv](https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv)

* Test Data

[https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv](https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv)


It was uploaded into the project as follows: 
```{r, cache=TRUE}
training_data <- read.csv("pml-training.csv")
test_data     <- read.csv("pml-testing.csv")
```


#Data Clean Up


To facilitate a predictive model, the data in the training and testing data sets must be common.

To ensure the above condition is met, fields that are not common in both the training set and the test set as DROPPED!!
```{r}
drops <- colnames(test_data[, ( colSums(is.na(test_data)) / nrow(test_data) ) > 0.8 ] )
training_data2 <- training_data[, !(names(training_data) %in%  drops     )]

training_data2 <- training_data2[, !( names(training_data2) ==   colnames(test_data[1]))  ]
training_data2 <- training_data2[, !( names(training_data2) ==   colnames(test_data[3:6]))  ]

```

If fields are highly correlated in the data set, there is a redundancy in having both fields present. Additionally, this redundancy may have the adverse effect of overfitting our model.

We remove correlated values (cor > 84% ), as follows
```{r}
num_fld <- sapply( training_data2 ,  is.numeric )
training_data3 <- training_data2[ , -findCorrelation(   cor(training_data2[,num_fld])  , cutoff=.9  ) ]

```


# Data Partition 

We partition the training set from approve to 75% training, and 25% testing.

```{r}

set.seed(1024)
inTrain = createDataPartition(training_data3$classe , p = 3/4)[[1]]
training = training_data3[ inTrain,]
testing = training_data3[-inTrain,]
```

# Model Training and Selection

We train our model using 

* Random Forest
* Recursive Partition

```{r, cache=TRUE}

set.seed(12345)
# Recursive Partition Tress
modelFitRP <- train( classe ~ ., method = "rpart", data = training )
#Random Forest
modelFitRF <- train( classe ~ ., method = "rf", data = training )

```

We then test our model based on the data we partitioned

```{r , message=FALSE}

predRP <- predict( modelFitRP , testing )
predRF <- predict( modelFitRF , testing )
```

##Confusion Matrix for Recursive Partition
```{r }
confusionMatrix( testing$classe , predRP  )
```

The recursive partition performs poorly. This is based off the Sensisitivity Values and the confusion matrix.


##Confusion Matrix for Random Forest
```{r}
confusionMatrix( testing$classe , predRF  )
```

The Random Forest perform comparatively better than Recursive Partition. This is based off the Sensisitivity Values and the confusion matrix.

WE SELECT RANDOM FOREST AS OUR ALGORITHIM.

#Predict Activitiy Using Test Data

We run the Random Forest model against the data ...
```{r , message=FALSE}
predRF_test_data <- predict( modelFitRF , test_data )
data.frame( test_data[,1:2], as.data.frame(predRF_test_data) )
```




